---
title: "Do A-League Defenders Tackle More Than Midfielders?"
author: "Untapped Talent: Australian Football Scouting & Analysis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

Conventional wisdom says tackling is a defender's job. This analysis tests that
assumption against real 2025/26 A-League Men player data, using position group
comparisons and formal hypothesis testing.

## Loading and Preparing the Data

```{r load-data}
library(readxl)
library(dplyr)
library(ggplot2)

standard <- read_excel("A_League_Stats_Cleaned.xlsx", sheet = "Players Standard Sheet")
defense <- read_excel("Defense_Passing_Cleaned.xlsx", sheet = "Defense Players stats")

merged <- standard %>%
  left_join(defense %>% select(Player, Squad, `Tackles Won`), by = c("Player", "Squad")) %>%
  mutate(PrimaryPos = sub(",.*", "", Position)) %>%
  filter(PrimaryPos != "GK")  # exclude goalkeepers - not a meaningful comparison group
```

## Sample Sizes by Position

```{r sample-sizes}
table(merged$PrimaryPos)
```

## Visualizing the Distribution

```{r boxplot}
ggplot(merged, aes(x = PrimaryPos, y = `Tackles Won`, fill = PrimaryPos)) +
  geom_boxplot(alpha = 0.8, outlier.shape = 21, outlier.size = 2.5, outlier.fill = "white") +
  geom_jitter(width = 0.15, alpha = 0.35, size = 1.5, color = "black") +
  scale_fill_manual(values = c("DF" = "#2E86AB", "MF" = "#F18F01", "FW" = "#C73E1D")) +
  labs(
    title = "Tackles Won by Position",
    subtitle = "A-League Men, 2025/26 season",
    x = "Position",
    y = "Tackles Won"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(color = "gray40"),
    panel.grid.minor = element_blank()
  )
```

The boxplot shows Defenders (DF) and Midfielders (MF) with visually similar
medians, while Forwards (FW) sit noticeably lower. Several high-tackling
outliers are visible in both DF and MF groups.

## Testing for Normality

Before choosing between ANOVA and a non-parametric alternative, we test
whether Tackles Won is normally distributed:

```{r shapiro}
shapiro.test(merged$`Tackles Won`)
```

The p-value is far below 0.05, indicating the data significantly deviates
from a normal distribution. This rules out a standard one-way ANOVA as the
appropriate test here, since ANOVA assumes normally distributed data within
each group.

## Kruskal-Wallis Test

Given the non-normal distribution, we use the Kruskal-Wallis test — the
non-parametric equivalent of a one-way ANOVA — to test whether Tackles Won
differs significantly across the three position groups.

```{r kruskal}
kruskal.test(`Tackles Won` ~ PrimaryPos, data = merged)
```

The result is highly significant (p < 0.001), indicating at least one
position group differs meaningfully from the others.

## Post-Hoc Pairwise Comparison (Dunn's Test)

Kruskal-Wallis tells us a difference exists somewhere, but not which specific
pairs of groups differ. We follow up with Dunn's test, using a Bonferroni
correction to account for running multiple pairwise comparisons.

```{r dunn}
library(FSA)
dunnTest(`Tackles Won` ~ PrimaryPos, data = merged, method = "bonferroni")
```

## Results Summary

```{r results-table, echo=FALSE}
results <- data.frame(
  Comparison = c("DF vs FW", "DF vs MF", "FW vs MF"),
  `Adjusted P-value` = c("8.21e-08", "1.00", "2.38e-08"),
  `Significant (a = 0.05)` = c("Yes", "No", "Yes"),
  check.names = FALSE
)
knitr::kable(results)
```

## Conclusion

Forwards tackle significantly less than both Defenders and Midfielders, which
is expected. However, **Defenders and Midfielders show no statistically
significant difference in Tackles Won** (p = 1.00, adjusted). Despite the
positional label, A-League midfielders in the 2025/26 season are contributing
a statistically indistinguishable volume of tackling compared to defenders.

This may reflect modern pressing systems that ask midfielders to contest the
ball higher up the pitch, rather than defensive responsibility sitting solely
with the back line.

## Limitations

- This analysis uses `PrimaryPos`, derived by taking the first listed position
  for players with multiple position tags (e.g. "MF,FW" becomes "MF"). Some
  players may be miscategorized if they play a genuinely hybrid role.
- Tackles Won alone doesn't capture full defensive contribution (e.g.
  interceptions, blocks, clearances). Deeper defensive metrics like
  Clearances were unavailable at the time of writing due to FBref's loss of
  its Opta data licence in January 2026.
- Season-aggregate data cannot account for tactical differences between
  clubs, which may explain some of the within-group variation.
