---
title: "Home Advantage in the A-League: Points Say No, Goals Say Yes"
subtitle: "Untapped Talent: Australian Football Scouting & Analysis"
date: "2026-09-23"
output: html_document
---

## Hypothesis

Is home advantage real in the A-League, or is it a myth? Home and away
records are paired by team (13 home games, 13 away games each), so this
uses paired tests rather than independent-sample tests.

*Data: Soccerway, 2025–26 A-League Men season, home/away splits.*

```{r}
df <- read.csv("aleague 202526 home away.csv")
```

## Points per game

```{r}
diff <- df$home_ppg - df$away_ppg

shapiro.test(diff)
```

Shapiro-Wilk is not significant (p = 0.241), so the differences are
normally distributed and a paired t-test is appropriate.

```{r}
t.test(df$home_ppg, df$away_ppg, paired = TRUE)
```

```{r}
mean(diff)
d_cohen_ppg <- mean(diff) / sd(diff)
d_cohen_ppg
```

Home teams average 0.096 more points per game than away teams, but this
is not statistically significant (p = 0.198). Cohen's d = 0.395 is a
small-to-medium effect, with only 12 teams, the test is underpowered
to detect it.

## Goal difference

```{r}
gd_diff <- df$home_gd - df$away_gd

shapiro.test(gd_diff)
```

Normal again (p = 0.351), so paired t-test applies.

```{r}
t.test(df$home_gd, df$away_gd, paired = TRUE)
```

```{r}
mean(gd_diff)
d_cohen_gd <- mean(gd_diff) / sd(gd_diff)
d_cohen_gd
```

Home teams average +2.42 goal difference, away teams -2.42 which is a swing of
4.83 per team, and this **is** statistically significant (p = 0.0064),
with a large effect size (d = 0.969).

## Conclusion

Points say home advantage isn't real in the A-League. Goal difference
says otherwise. Points are a blunt instrument a 1-0 win and a 4-0 win
count the same, so home advantage shows up in how convincingly teams
win or lose, even when it doesn't reliably convert into extra points.


